Incremental tactics

### 2012-05-11 doc/papers/LFMTP2012/paper.tex
paper + new Test_check3

>>>>

A student is learning about System \sysname T, and wants to write a
program to compute $4!$. He (or she) knows that System \sysname T is a
typed language, and wants to take advantage of this to build his
program incrementally using the teacher's provided incremental
verifier. As a first attempt, he writes the following program $P_1$:
$$
\recb{\s\z}{\s\z}{\lam x\lam y \s {\var x}}
$$
The system type-checks it correctly by constructing a typing
derivation $\md_1$ of \/ $\vdash P_1 : \cst{nat}$, valid in an
empty context, and names all subderivations constructed for future
reuse. The student then decides to factorize a definition for
$\var{add}$ out of his code. He writes:
\begin{align*}
  &\letb{add}{\tlam x {\cst{nat}} \tlam y {\cst{nat}} \recb {\var x}
    {\var y} {\lam x \lam y \s{\var x}}}
  \\
  &\quad\app{\var{add}}\app{(\s\z)}{(\s\z)} \ \text{.}
\end{align*}
But he (or his text editor) realizes that some parts of this program
have not changed: both $\s\z$, and the function $\lam x\lam y\s{\var
  x}$. The system must then have a name for the already-built
associated derivations $\md_2$, $\md_3$ and $\md_4$. Provided there is
an operator $|\md|$ from the derivation to the program it types, what
is really to be checked is:
\begin{align*}
  &\letb{add}{\tlam x {\cst{nat}} \tlam y {\cst{nat}} \recb {\var x}
    {\var y} {|\md_4|}}
  \app{\var{add}}\app{|\md_2|}{|\md_2|} \ \text{,}
\end{align*}
that is, only the subterms of the program that changed. Derivations
$\md_2$ and $\md_4$ used to live in the empty context, we are then
free to use it in the new context where
$\var{add}:\cst{nat}\to\cst{nat}$, because System \sysname T enjoys
the \emph{weakening} property. So the system answers with a new
derivation made out of pieces of the old one, without having to
recheck the already built ones. It also gives the name $\md_5$ to the
derivation for $\tlam x {\cst{nat}} \tlam y {\cst{nat}} \recb {\var x}
{\var y} {|\md_4|}$, $\md_6$ to the arguments $\app{|\md_2|}{|\md_2|}$
and $\md_7$ to the call $\app{\var{add}} {|\md_6|}$. Even if it
occured in the scope of $\var{add}$, $\md_6$ doesn't depend on the
fact that $\vdash \var{add}:\cst{nat}\to\cst{nat}$, but
$\md_7$ does: %TODO
. The
student then types:
\begin{align*}
  &\letb{add}{ |\md_5| }
  \\
  &\letb{mul}{ \tlam x {\cst{nat}} \tlam y
    {\cst{nat}} \recb \z {\var y} {\lam z \lam t
      \app{\var{add}}\app{\var x} {\var z}} }
  \\
  &\app{\var{mul}}\app{|\md_2|}{|\md_2|}
\end{align*}
to see if his definition of multiplication is well-typed.
